---
title: "Monte Carlo"
author: "Broderick Prows"
date: "1/30/2020"
output: html_document
---

Estimating $\pi$
```{r}
pi
```


```{r}
sample_size <- 1
x <- runif(sample_size)
y <- runif(sample_size)
4*mean(x**2 + y**2 <= 1)
```

Can also use Monte Carlo to estimate percentiles. Suppose X ∼ N(200, 182). 
(`r qnorm(.80, mean=200, sd=18)`)
```{r}
sim_vals <- rnorm(15000,mean = 200, sd = 18)
quantile(sim_vals, .80)
```

Theta1 and Theta2, Comparing the differences
```{r}
theta1s <- rbeta(6000, 75, 123)
theta2s <- rbeta(6000, 91, 53)
diffs = theta1s - theta2s
mean(diffs)
```

95% Credible interval
```{r}
quantile(diffs, c(0.025, 0.975))
```
Since the interval does not catch 0, there is a statistically significant difference. 

Estimated Posterior Probability that theta2 exceeds theta1 by at least 0.2
```{r}
mean(diffs <= -0.2)
```

Density Plot
```{r}
t1s <- rbeta(1000000, 75,123)
t2s <- rbeta(1000000,91,53)
plot(density(t1s-t2s),
xlab=expression(theta[1]-theta[2]),
main=expression(paste("Estimated Posterior Density of ",
theta[1], "-", theta[2]), sep=""))
```

Home Field Advantage
`r expression(theta)` is defined as the probability that the home team wins in a randomly selected NFL Game.
Prior: $$\theta = Beta(a = 5, b = 4)$$

Data: n = 260, y = 136
```{r}
a = 5
b = 4
n = 260
y = 136
a_star = a + y
b_star = b + n - y
thetas <- seq(0,1,length.out=1001)
interval <- c(qbeta(.05,shape1=a_star,shape2=b_star), qbeta(.95,shape1=a_star,shape2=b_star))
interval
plot(thetas, dbeta(thetas, shape1 = a_star, shape2 = b_star), type = 'l', col = 'black',
     main = 'Home Field Advantage', xlab = expression(theta), ylab = 'Density')
lines(thetas, dbeta(thetas, shape1 = 5, shape2 = 4), type = 'l', col = 'gray')
legend("topright", legend=c("Posterior", "Prior"),
       col=c("black", "gray"), lty = 1, lwd = 2, cex = 1)
abline(v=interval,col='darkgrey',lwd=2)
text(interval[1]-.05, dbeta(interval[1],a_star,b_star), round(interval[1], 3))
text(interval[2]+.05, dbeta(interval[2],a_star,b_star), round(interval[2], 3))
```

```{r}
1 - pbeta(.50, shape1 = a_star, shape2 = b_star)
```



