---
title: "HW3"
author: "Broderick Prows"
date: "1/16/2020"
output: html_document
---

Questions:
1. Do you support change in Honor Code to allow beards?
2. Do you support capital punishment?
3. Speak a foreign language?
4. Do you support Trump impeachment?
5. Can you drive a stick-shift?
6. Ever traveled outside of the country?


Using:
1. Do you support change in Honor Code to allow beards?
Priors:
```{r}
options(scipen=999)
theta <- c(.25, .5, .75)
p <- c(1/3, 1/3, 1/3)

bayes_update_bernoli <- function(theta, priors, n, y) {
  lkhd <- (theta ** y) * ( (1 - theta) ** (n - y)); lkhd
  part_marg <- priors * lkhd; part_marg
  marg_y <- sum(part_marg); marg_y
  priors <- part_marg / marg_y;
  bundle <- list("priors" = priors,
                 "LKHD" = lkhd,
                 "mid_marg" = part_marg,
                 "Marginal" = marg_y)
  return(bundle)
}
```

  {y1 = 0}
Bayes Update
```{r}
first <- bayes_update_bernoli(theta, p, n = 1, y = 0)
first
```


{y2 = 1, y3 = 1, y4 = 1, y5 = 1, y6 = 1}
```{r}
then <- bayes_update_bernoli(theta, first$priors, n = 5, y = 5)
then
```

{y7 = 1, y8 = 1, y9 = 1, y10 = 1, y11 = 0}
```{r}
last <- bayes_update_bernoli(theta, then$priors, n = 5, y = 4)
last
```

What about all at once?
```{r}
bayes_update_bernoli(theta, p, n = 11, y = 9)
```

Using the binomial summary statistic.
```{r}
bayes_update_binomial <- function(theta, priors, n, y) {
  lkhd <- choose(n,y) * (theta ** y) * ( (1 - theta) ** (n - y))
  part_marg <- priors * lkhd
  marg_y <- sum(part_marg)
  priors <- part_marg / marg_y
  bundle <- list("priors" = priors,
                 "LKHD" = lkhd,
                 "mid_marg" = part_marg,
                 "Marginal" = marg_y)
  return(bundle)
}
```


```{r}
bayes_update_binomial(theta, p, n = 11, y = 9)
```



